---
title: "CellChat Analysis Report"
params:
  file1: NULL
  file2: NULL
  sample1: NULL
  sample2: NULL
output:
  html_document:
    theme:
      version: 5
      bootswatch: lumen
    toc: true
    toc_depth: 2
runtime: shiny
---
<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
library(CellChat)
library(patchwork)
library(downloadthis)
library(Cairo)
options(shiny.usecairo = TRUE)
```

# Report {.tabset}

## Single sample analysis

### Introduction

This report presents a comprehensive analysis of cell–cell communication using the CellChat package. Each section includes visualizations and tables to illustrate key interaction patterns and signaling changes between groups.


```{r load-cellchat, echo=FALSE, results='hide'}
cellchat1 <- readRDS("ctrl_cellchat.rds")

cellchat1 <- netAnalysis_computeCentrality(cellchat1, slot.name = "netP")

pathway_list <- cellchat1@netP$pathways

# cellchat <- readRDS("merged_cellchat.rds")
```

### Signaling role 
```{r select-pathways, fig.width=6, fig.height=5}
fluidRow(
  # Left column: pathway selector + download button
  column(
    width = 2,
    selectInput(
      inputId  = "pathway",
      label    = "Select a signalling pathway",
      choices  = pathway_list,
      selected = pathway_list[1],
      multiple = FALSE
    ),
    br(),
    downloadButton("download_signaling_role_pdf", "Download PDF"),
    br(),br(),
    downloadButton("download_signaling_role_svg", "Download SVG")
  ),

  # Right column: the network plot
  column(
    width = 10,
    plotOutput("signaling_role_plot", height = "500px")
  )
)

# selectInput(
#   inputId  = "pathway",
#   label    = "Select a signalling pathway",
#   choices  = pathway_list,
#   selected = pathway_list[1],
#   multiple = FALSE
# )
# 
# renderPlot({
#   netAnalysis_signalingRole_network(cellchat, signaling = input$pathway, width = 30, height = 10, font.size = 11, font.size.title = 12)
# })

br()
br()

```

```{r sig-network}

output$signaling_role_plot <- renderPlot({
  # req(input$pathway)
  netAnalysis_signalingRole_network(
    cellchat1,
    signaling       = input$pathway,
    width           = 30,
    height          = 10,
    font.size       = 11,
    font.size.title = 12
  )
})

# 2) download handler to save the network as a single‐page PDF
output$download_signaling_role_pdf <- downloadHandler(
  filename = function() {
    paste0("signaling_role_network_", Sys.Date(), ".pdf")
  },
  content = function(file) {
    # req(input$pathway)
    # open PDF device
    pdf(file, width = 14, height = 6)
    # capture the plot object
    p <- netAnalysis_signalingRole_network(
      cellchat1,
      signaling       = input$pathway,
      width           = 30,
      height          = 10,
      font.size       = 11,
      font.size.title = 12
    )
    # force rendering
    print(p)
    dev.off()
  },
  contentType = "application/pdf"
)

output$download_signaling_role_svg <- downloadHandler(
  filename = function() {
    paste0("signaling_role_network_", Sys.Date(), ".svg")
  },
  content = function(file) {
    # req(input$pathway)
    # open PDF device
    svg(file, width = 14, height = 6)
    # capture the plot object
    p <- netAnalysis_signalingRole_network(
      cellchat1,
      signaling       = input$pathway,
      width           = 30,
      height          = 10,
      font.size       = 11,
      font.size.title = 12
    )
    # force rendering
    print(p)
    dev.off()
  },
  contentType = "image/svg+xml"
)

```
### Bubble plot
```{r cell-selection}

fluidRow(
  # Left column: inputs + download button
  column(
    width = 2,
    selectInput(
      inputId  = "pathway1",
      label    = "Select a signalling pathway",
      choices  = pathway_list,
      selected = pathway_list[1],
      multiple = FALSE
    ),
    selectInput(
      inputId  = "source_cells",
      label    = "Select source cell groups",
      choices  = levels(cellchat1@idents),
      selected = NULL,
      multiple = TRUE
    ),
    selectInput(
      inputId  = "target_cells",
      label    = "Select target cell groups",
      choices  = levels(cellchat1@idents),
      selected = NULL,
      multiple = TRUE
    ), br(),
    downloadButton("download_bubble_pdf", "Download PDF"),
    br(),br(),
    downloadButton("download_bubble_svg", "Download SVG")
  ),

  # Right column: the bubble plot itself
  column(
    width = 10,
    plotOutput("bubble_plot", height = "400px", width = "1000px")
  )
)
```


```{r bubble-plot, fig.width=6, fig.height=5}


# downloadButton("download_bubble", "Download Bubble Plot as PDF")

output$bubble_plot <- renderPlot({
  #req(input$pathway, input$source_cells, input$target_cells)
  netVisual_bubble(
    cellchat1,
    signaling      = input$pathway1,
    sources.use    = input$source_cells,
    targets.use    = input$target_cells,
    remove.isolate = FALSE,
    dot.size.min   = 2,
    dot.size.max   = 12,
    font.size      = 14,
    font.size.title= 14,
    line.size      = 0.4,
    angle.x        = 45
  )
})

# 2) download handler to produce a single‐page PDF of the bubble plot
output$download_bubble_pdf <- downloadHandler(
  filename = function() {
    paste0("bubble_plot_", Sys.Date(), ".pdf")
  },
  content = function(file) {
    # req(input$pathway, input$source_cells, input$target_cells)
    # 1) Open a PDF device
    pdf(file, width = 15, height = 5)
    # 2) Capture the plot object
    p <- netVisual_bubble(
      cellchat1,
      signaling      = input$pathway1,
      sources.use    = input$source_cells,
      targets.use    = input$target_cells,
      remove.isolate = FALSE,
      dot.size.min   = 2,
      dot.size.max   = 12,
      font.size      = 14,
      font.size.title= 14,
      line.size      = 0.4,
      angle.x        = 45
    )
    # 3) Force it onto the PDF
    print(p)
    # 4) Close device
    dev.off()
  },
  contentType = "application/pdf"
)

output$download_bubble_svg <- downloadHandler(
  filename = function() {
    paste0("bubble_plot_", Sys.Date(), ".svg")
  },
  content = function(file) {
    # req(input$pathway, input$source_cells, input$target_cells)
    # 1) Open a PDF device
    svg(file, width = 15, height = 5)
    # 2) Capture the plot object
    p <- netVisual_bubble(
      cellchat1,
      signaling      = input$pathway1,
      sources.use    = input$source_cells,
      targets.use    = input$target_cells,
      remove.isolate = FALSE,
      dot.size.min   = 2,
      dot.size.max   = 12,
      font.size      = 14,
      font.size.title= 14,
      line.size      = 0.4,
      angle.x        = 45
    )
    # 3) Force it onto the PDF
    print(p)
    # 4) Close device
    dev.off()
  },
  contentType = "image/svg+xml"
)
# renderPlot({
#   netVisual_bubble(
#     cellchat,
#     signaling     = input$pathway,
#     sources.use   = input$source_cells,
#     targets.use   = input$target_cells,
#     remove.isolate = FALSE,
#     dot.size.min   = 2,     # optional – keep very small bubbles visible
#     dot.size.max   = 12,     # ← bigger bubbles
#     font.size = 14,
#     font.size.title = 14,
#     line.size = 0.4,
#     angle.x = 45
#   )
# })


```

<!-- # bubble -->
<!-- ```{r} -->
<!--   netVisual_bubble( -->
<!--     cellchat, -->
<!--     signaling     = "ANNEXIN", -->
<!--     sources.use   = c("CD14 Mono", "CD4 Naive T"), -->
<!--     targets.use   = "DC", -->
<!--     remove.isolate = FALSE, -->
<!--     dot.size.min   = 2,     # optional – keep very small bubbles visible -->
<!--     dot.size.max   = 12,     # ← bigger bubbles -->
<!--     font.size = 14, -->
<!--     font.size.title = 14, -->
<!--     line.size = 0.4 -->
<!--   ) -->

<!-- ``` -->

### Circle plot

```{r select-pathways2, fig.width=6, fig.height=5}

selectInput(
  inputId  = "pathway_circle",
  label    = "Select a signalling pathway",
  choices  = pathway_list,
  selected = pathway_list[1],
  multiple = TRUE
)

downloadButton("download_plots_pdf", "Download All Plots as PDF")
downloadButton("download_plots_svg", "Download All Plots as SVG")

```

```{r, circle-plot, fig.width=6, fig.height=5}
# renderPlot({
#   netVisual_aggregate(cellchat, signaling = input$pathway, vertex.receiver = vertex.receiver, layout = "circle",
#                       vertex.weight = 3,
#                       vertex.size.max = 5,
#                       vertex.weight.max = 5,
#                       title.space = 4,
#                       alpha.image = 0.5,
#                       pt.title = 14,
#                       point.size = 3
#                       )
# })

chunk <- function(x, n) {
  split(x, ceiling(seq_along(x)/n))
}

output$plots_ui <- renderUI({
  req(input$pathway_circle)
  # split selected pathways into rows of max 3
  rows <- chunk(input$pathway_circle, 3)
  
  # build one fluidRow per chunk
  tagList(lapply(rows, function(path_vec) {
    fluidRow(
      lapply(path_vec, function(p) {
        column(
          width = 4,       # 12/3 = 4 columns each
          plotOutput(
            outputId = paste0("plot_", make.names(p)),
            height   = "350px"  # adjust as you like
          )
        )
      })
    )
  }))
})

# render each plot on the fly
observe({
  req(input$pathway_circle)
  lapply(input$pathway_circle, function(p) {
    out_id <- paste0("plot_", make.names(p))
    output[[out_id]] <- renderPlot({
      netVisual_aggregate(
        cellchat1,
        signaling         = p,
        vertex.receiver   = vertex.receiver,
        layout            = "circle",
        vertex.weight     = 3,
        vertex.size.max   = 5,
        vertex.weight.max = 5,
        title.space       = 4,
        alpha.image       = 0.5,
        pt.title          = 14,
        point.size        = 3
      )
    })
  })
})

uiOutput("plots_ui")

output$download_plots_pdf <- downloadHandler(
  filename = function() {
    paste0("circle_plots_", Sys.Date(), ".pdf")
  },
  content = function(file) {
    req(input$pathway_circle)
    # open PDF device
    pdf(file, width = 6, height = 5)
    # draw one page per pathway
    for (p in input$pathway_circle) {
      netVisual_aggregate(
        cellchat1,
        signaling         = p,
        vertex.receiver   = vertex.receiver,
        layout            = "circle",
        vertex.weight     = 3,
        vertex.size.max   = 5,
        vertex.weight.max = 5,
        title.space       = 4,
        alpha.image       = 0.5,
        pt.title          = 14,
        point.size        = 3
      )
    }
    dev.off()
  },
  contentType = "application/pdf"
)

# output$download_plots_svg <- downloadHandler(
#   filename = function() {
#     paste0("circle_plots_", Sys.Date(), ".svg")
#   },
#   content = function(file) {
#     req(input$pathway_circle)
#     pathways  <- input$pathway_circle
#     n         <- length(pathways)
#     ncol      <- 3
#     nrow      <- ceiling(n / ncol)
#     
#     svg(file, width = ncol * 4, height = nrow * 4)
#     
#     grid::grid.newpage()
#     grid::pushViewport(
#       grid::viewport(
#         layout = grid::grid.layout(nrow, ncol)
#       )
#     )
#     for (i in seq_along(pathways)) {
#       p      <- pathways[i]
#       row    <- ceiling(i / ncol)
#       col    <- ((i - 1) %% ncol) + 1
#       
#       # drill into the right cell
#       grid::pushViewport(
#         grid::viewport(
#           layout.pos.row = row,
#           layout.pos.col = col
#         )
#       )
#       
#       # draw the plot
#       netVisual_aggregate(
#         cellchat,
#         signaling         = p,
#         vertex.receiver   = vertex.receiver,
#         layout            = "circle",
#         vertex.weight     = 3,
#         vertex.size.max   = 5,
#         vertex.weight.max = 5,
#         title.space       = 4,
#         alpha.image       = 0.5,
#         pt.title          = 14,
#         point.size        = 3
#       )
#       
#       grid::popViewport()  # back out of this cell
#     }
# 
#     grid::popViewport()
#     dev.off()
#   },
#   contentType = "image/svg+xml"
# )


output$download_plots_svg <- downloadHandler(
  filename   = function() paste0("circle_plots_", Sys.Date(), ".svg"),
  contentType= "image/svg+xml",
  content    = function(file) {
    req(input$pathway_circle)
    paths <- input$pathway_circle
    n     <- length(paths)
    
    # decide rows & columns
    ncol <- min(3, n)
    nrow <- ceiling(n / ncol)
    
    # 1) open SVG device
    svg(file,
        width  = ncol * 5,   # 5 inches per panel horizontally
        height = nrow * 7)   # 5 inches per panel vertically
    
    # 2) tell base graphics to split the page
    par(
      mfrow = c(nrow, ncol),
      mar   = c(1,1,2,1),   # tighten margins (bottom, left, top, right)
      xpd   = TRUE
    )
     
    # 3) draw each circle plot
    for (p in paths) {
      netVisual_aggregate(
        cellchat1,
        signaling         = p,
        vertex.receiver   = vertex.receiver,
        layout            = "circle",
        vertex.weight     = 3,
        vertex.size.max   = 5,
        vertex.weight.max = 5,
        title.space       = 4,
        alpha.image       = 0.5,
        pt.title          = 14,
        point.size        = 3
      )
    }
    
    # 4) close device
    dev.off()
  }
)

```


## Comparison Analysis

### Introduction

This report presents a comprehensive analysis of cell–cell communication using the CellChat package. Each section includes visualizations and tables to illustrate key interaction patterns and signaling changes between groups.

```{r load-cellchat2, echo=FALSE, results='hide'}
cellchat.NL <- readRDS("ctrl_cellchat.rds")
cellchat.LS <- readRDS("stim_cellchat.rds")

s1 <- "ctrl"
s2 <- "stim"

object.list1 <- setNames(list(cellchat.NL, cellchat.LS), c(s1, s2))
cellchat <- mergeCellChat(object.list1, add.names = names(object.list1), cell.prefix = TRUE)


cellchat.NL<-netAnalysis_computeCentrality(cellchat.NL)
cellchat.LS<-netAnalysis_computeCentrality(cellchat.LS)

# cellchat <- readRDS("merged_cellchat.rds")
```

<!-- # Interaction Comparison {.tabset} -->

<!-- We compare the number of interactions and the interaction weights between two experimental groups. -->

<!-- ## Total number of interactions -->

<!-- ```{r compare-count} -->

<!-- selectInput( -->

<!--   'measure', label = 'Number of bins:', -->

<!--   choices = c("count","weight"), selected = "count" -->

<!-- ) -->

<!-- renderPlot({ -->

<!--   compareInteractions(cellchat, show.legend = FALSE, group = c(1, 2), measure = input$measure) -->

<!-- }) -->

<!-- #plot(gg1) -->

<!-- ``` -->

### Total number of interactions

<div style="width:100%; display:flex; justify-content:space-between; align-items:flex-start;">
  <div style="width:48%; text-align:center;">
    
```{r echo=FALSE, fig.width=6, fig.height=5}
    gg1 <- compareInteractions(cellchat, show.legend=FALSE, group=c(1,2))
    gg1
```

```{r echo=FALSE}
download_this(
  gg1,
  output_name = "cellchat_interactions",
  output_extension = ".pdf",
  button_label = "Download PDF",
  button_type = "default",
  self_contained = FALSE,
  ggsave_args = list(width = 8, height = 6, dpi = 300)
)

download_this(
  gg1,
  output_name = "cellchat_interactions",
  output_extension = ".svg",
  button_label = "Download SVG",
  button_type = "default",
  self_contained = FALSE,
  ggsave_args = list(width = 8, height = 6, dpi = 300)
)
```
  </div>

  <div style="width:48%; text-align:center;">
```{r echo=FALSE, fig.width=6, fig.height=5}
gg2 <- compareInteractions(cellchat, show.legend=FALSE, group=c(1,2), measure="weight")
print(gg2)
```

```{r echo=FALSE}
download_this(
  gg2,
  output_name = "cellchat_interactions_weight",
  output_extension = ".pdf",
  button_label = "Download PDF",
  button_type = "default",
  has_icon = TRUE,
  icon = "fa fa-save"
)

download_this(
  gg2,
  output_name = "cellchat_interactions_weight",
  output_extension = ".svg",
  button_label = "Download SVG",
  button_type = "default",
  has_icon = TRUE,
  icon = "fa fa-save"
)
```
  </div>
</div>

### Number of interactions and interaction strength among different cell populations {.tabset}

This section visualizes differences in network structure between groups using raw counts and weighted measures.

#### Circle Plot across two datasets

<div style="width:100%; display:flex; justify-content:space-between; align-items:flex-start;">
  <div style="width:48%; text-align:center;">

```{r diff-count, fig.width=6, fig.height=5}
p1 <- netVisual_diffInteraction(cellchat, weight.scale = TRUE, measure = "count", margin = 0)
print(p1)
```

```{r echo=FALSE}
tmp_pdf <- "circleplot_interaction_count.pdf"
pdf(tmp_pdf,
    width  = 8,   # 8" × 300 dpi
    height = 7
    )
replayPlot(p1)
invisible(dev.off())

download_file(
  path           = "circleplot_interaction_count.pdf",
  output_name      = "circleplot_interaction_count",
  button_label     = "Download PDF",
  button_type      = "default",
  self_contained   = FALSE
)

tmp_svg <- "circleplot_interaction_count.svg"
svg(tmp_svg,
    width  = 8,   # 8" × 300 dpi
    height = 7)
replayPlot(p1)
invisible(dev.off())

download_file(
  path           = "circleplot_interaction_count.svg",
  output_name      = "circleplot_interaction_count",
  button_label     = "Download SVG",
  button_type      = "default",
  self_contained   = FALSE
)

```
  </div>

  <div style="width:48%; text-align:center;">
```{r echo=FALSE, fig.width=6, fig.height=5}
p2 <- netVisual_diffInteraction(cellchat, weight.scale = TRUE, measure = "weight", margin = 0)
print(p2)
```

```{r echo=FALSE}
tmp_pdf2 <- "circleplot_interaction_weight.pdf"
pdf(tmp_pdf2,
    width  = 8 ,   # 8" × 300 dpi
    height = 7)
replayPlot(p2)
invisible(dev.off())

download_file(
  path           = "circleplot_interaction_weight.pdf",
  output_name      = "circleplot_interaction_weight",
  button_label     = "Download PDF",
  button_type      = "default",
  self_contained   = FALSE
)

tmp_svg2 <- "circleplot_interaction_weight.svg"
svg(tmp_svg2,
    width  = 8,   # 8" × 300 dpi
    height = 7)
replayPlot(p2)
invisible(dev.off())

download_file(
  path           = "circleplot_interaction_weight.svg",
  output_name      = "circleplot_interaction_weight",
  button_label     = "Download SVG",
  button_type      = "default",
  self_contained   = FALSE
)

```
  </div>
</div>

#### Heatmap across two datasets
We generate circle plots of interaction networks for each sample in the provided list. The dot sizes are scaled by the maximum interaction weight.

<div style="width:100%; display:flex; justify-content:space-between; align-items:flex-start;">
  <div style="width:48%; text-align:center;">

```{r diff-heatmap, fig.width=6, fig.height=5}
library(ComplexHeatmap)

h1 <- netVisual_heatmap(cellchat, measure = "count")
print(h1)
pdf("diff_heatmap_count.pdf",
    width  = 7,   # 8" × 300 dpi
    height = 6)      # res    = 300
ComplexHeatmap::draw(h1)
invisible(dev.off())

svg("diff_heatmap_count.svg",
    width  = 7,   # 8" × 300 dpi
    height = 6)
ComplexHeatmap::draw(h1)
invisible(dev.off())
```

```{r echo=FALSE}
download_file(
  path        = "diff_heatmap_count.pdf",
  output_name = "diff_heatmap_count",
  button_label = "Download\nPDF",
  button_type = "default",
  self_contained = FALSE
)

download_file(
  path        = "diff_heatmap_count.svg",
  output_name = "diff_heatmap_count",
  button_label = "Download \n SVG",
  button_type = "default",
  self_contained = FALSE
)
```
  </div>

  <div style="width:48%; text-align:center;">
```{r diff-heatmap-weight, fig.width=6, fig.height=5}
h2 <- netVisual_heatmap(cellchat, measure = "weight")
print(h2)
# png("diff_heatmap_weight.png",
#     width  = 8 * 300,   # 8" × 300 dpi
#     height = 6 * 300,
#     res    = 300)
# ComplexHeatmap::draw(h2)
# invisible(dev.off())

pdf("diff_heatmap_weight.pdf",
    width  = 8,   # 8" × 300 dpi
    height = 7)
ComplexHeatmap::draw(h2)
invisible(dev.off())

svg("diff_heatmap_weight.svg",
    width  = 8,   # 8" × 300 dpi
    height = 7)
ComplexHeatmap::draw(h2)
invisible(dev.off())

```

```{r echo=FALSE}
download_file(
  path        = "diff_heatmap_weight.pdf",
  output_name = "diff_heatmap_weight",
  button_label = "Download \n PDF",
  button_type = "default",
  self_contained = FALSE
)

download_file(
  path        = "diff_heatmap_weight.svg",
  output_name = "diff_heatmap_weight",
  button_label = "Download \n SVG",
  button_type = "default",
  self_contained = FALSE
)

```
  </div>
</div>

#### Between experimental groups
 
<div style="width:100%; display:flex; justify-content:space-between; align-items:flex-start;">
  <div style="width:48%; text-align:center;">
  
```{r circle-plot-1, fig.width=6, fig.height=5}
object.list <- setNames(list(cellchat.NL, cellchat.LS), c(s1, s2))
# compute max weight once
weight.max <- getMaxWeight(object.list, attribute = c("idents", "count"))

# render & save first plot

nc1<-netVisual_circle(
  object.list[[1]]@net$count,
  weight.scale    = TRUE,
  label.edge      = FALSE,
  edge.weight.max = weight.max[2],
  edge.width.max  = 12,
  title.name      = paste0("Number of interactions – ", names(object.list)[1])
)
# png("circle_plot_1.png",
#     width  = 8 * 300,
#     height = 6 * 300,
#     res    = 300)
# 
# replayPlot(nc1)
# 
# invisible(dev.off())

pdf("circle_plot_1.pdf",
    width  = 8,
    height = 7)

replayPlot(nc1)

invisible(dev.off())

svg("circle_plot_1.svg",
    width  = 8,
    height = 7)

replayPlot(nc1)

invisible(dev.off())
```

```{r echo=FALSE}
download_file(
  path           = "circle_plot_1.pdf",
  output_name    = "circle_plot_1",
  button_label   = "Download PDF",
  button_type    = "default",
  self_contained = FALSE
)

download_file(
  path           = "circle_plot_1.svg",
  output_name    = "circle_plot_1",
  button_label   = "Download SVG",
  button_type    = "default",
  self_contained = FALSE
)
```
  </div>
  
  <div style="width:48%; text-align:center;">
```{r circle-plot-2, fig.width=6, fig.height=5}
# reuse same max weight
weight.max <- getMaxWeight(object.list, attribute = c("idents", "count"))

# render & save second plot
nc2<-netVisual_circle(
  object.list[[2]]@net$count,
  weight.scale    = TRUE,
  label.edge      = FALSE,
  edge.weight.max = weight.max[2],
  edge.width.max  = 12,
  title.name      = paste0("Number of interactions – ", names(object.list)[2])
)
# png("circle_plot_2.png",
#     width  = 8 * 300,
#     height = 6 * 300,
#     res    = 300)
# 
# replayPlot(nc2)
# 
# invisible(dev.off())

pdf("circle_plot_2.pdf",
    width  = 8,
    height = 7)

replayPlot(nc2)

invisible(dev.off())

svg("circle_plot_2.svg",
    width  = 8,
    height = 7)

replayPlot(nc2)

invisible(dev.off())
```

```{r echo=FALSE}
download_file(
  path           = "circle_plot_2.pdf",
  output_name    = "circle_plot_2",
  button_label   = "Download PDF",
  button_type    = "default",
  self_contained = FALSE
)

download_file(
  path           = "circle_plot_2.svg",
  output_name    = "circle_plot_2",
  button_label   = "Download SVG",
  button_type    = "default",
  self_contained = FALSE
)
```

  </div>
</div>


#### Circle plots for selected cell types {.tabset}

```{r selectinput}

selectizeInput(
  inputId  = "levels",
  label    = "Cell types (select up to three):",
  choices  = levels(cellchat@idents$joint),
  selected = levels(cellchat@idents$joint)[1:3],
  multiple = TRUE,
  options  = list(maxItems = 3)
)

```

##### Between experimental groups

We merge interactions by major cell-type groups and visualize combined networks.

```{r data-process}

tagList(
  div(class = "row",
    lapply(seq_along(object.list), function(i) {
      div(class = "col-md-6",
        # Plot output
        plotOutput(outputId = paste0("mergedPlot_", i), height = "400px"),
        # Small spacer
        tags$br(),
        # Download button
        div(class = "text-center",
          downloadButton(
            outputId = paste0("download_pdf_", i),
            label    = "Download PDF"
          ),
          downloadButton(
            outputId = paste0("download_svg_", i),
            label    = "Download SVG"
          )
        ),
        # Add some bottom margin so rows don't stick together
        style = "margin-bottom: 30px;"
      )
    })
  )
)

```


```{r vis}
grp <- reactive({
  req(input$levels)
  factor(rep(input$levels, each = 4), levels = input$levels)
})
mergedList <- reactive({
  lapply(object.list, mergeInteractions, grp())
})
wmax <- reactive({
  getMaxWeight(
    mergedList(),
    slot.name   = c("idents", "net", "net"),
    attribute   = c("idents", "count", "count.merged")
  )
})

# 2) Loop to auto‐wire each plot + download
lapply(seq_along(object.list), function(i) {
  local({
    idx <- i
    nm  <- names(object.list)[idx]

    # render the circle plot
    output[[paste0("mergedPlot_", idx)]] <- renderPlot({
      netVisual_circle(
        mergedList()[[idx]]@net$count.merged,
        weight.scale     = TRUE,
        label.edge       = TRUE,
        edge.weight.max  = wmax()[3],
        edge.width.max   = 12,
        title.name       = paste0("Merged interactions – ", nm)
      )
    })

    # download handler for that same plot
    output[[paste0("download_pdf_", idx)]] <- downloadHandler(
      filename = function() {
        paste0("circle_plot_", nm, "_", Sys.Date(), ".pdf")
      },
      content = function(file) {
        pdf(file, width = 6, height = 5)
        netVisual_circle(
          mergedList()[[idx]]@net$count.merged,
          weight.scale     = TRUE,
          label.edge       = TRUE,
          edge.weight.max  = wmax()[3],
          edge.width.max   = 12,
          title.name       = paste0("Merged interactions – ", nm)
        )
        dev.off()
      },
      contentType = "application/pdf"
    )
    
    output[[paste0("download_svg_", idx)]] <- downloadHandler(
      filename = function() {
        paste0("circle_plot_", nm, "_", Sys.Date(), ".svg")
      },
      content = function(file) {
        svg(file, width = 6, height = 5)
        netVisual_circle(
          mergedList()[[idx]]@net$count.merged,
          weight.scale     = TRUE,
          label.edge       = TRUE,
          edge.weight.max  = wmax()[3],
          edge.width.max   = 12,
          title.name       = paste0("Merged interactions – ", nm)
        )
        dev.off()
      },
      contentType = NA
    )
  })
})

```

```{r merge-interactions}
# mergedCellchat <- reactiveVal(NULL)
# 
# plotOutput("mergedPlots", height = "400px")
# 
# # --- Server logic ----
# output$mergedPlots <- renderPlot({
#   # ensure inputs are available
#   req(input$levels, 4)
#   
#   # build the grouping factor
#   grp <- factor(
#     rep(input$levels, each = 4),
#     levels = input$levels
#   )
#   
#   # merge interactions on the fly
#   merged.list <- lapply(object.list, function(x) {
#     mergeInteractions(x, grp)
#   })
#   
#   # (re‐)build the combined CellChat if you need it later
#   cellchat.m <<- mergeCellChat(merged.list, add.names = names(merged.list))
#   mergedCellchat(cellchat.m)
#   # compute the max weight for scaling
#   wmax <- getMaxWeight(
#     merged.list,
#     slot.name = c("idents","net","net"),
#     attribute = c("idents","count","count.merged")
#   )
#   
#   netVisual_circle(
#       merged.list[[1]]@net$count.merged,
#       weight.scale     = TRUE,
#       label.edge       = TRUE,
#       edge.weight.max  = wmax[3],
#       edge.width.max   = 12,
#       title.name       = paste0("Merged interactions – ", names(merged.list)[i])
#   )
#   
#   
# })
# 
# downloadButton("download_plots_pdf", "Download All Plots as PDF")
# 
# output$download_plots_pdf <- downloadHandler(
#   filename = function() {
#     paste0("circle_plots_", Sys.Date(), ".pdf")
#   },
#   content = function(file) {
#     req(input$pathway_circle)
#     # open PDF device
#     pdf(file, width = 6, height = 5)
#     # draw one page per pathway
#     netVisual_circle(
#         merged.list[[1]]@net$count.merged,
#         weight.scale     = TRUE,
#         label.edge       = TRUE,
#         edge.weight.max  = wmax[3],
#         edge.width.max   = 12,
#         title.name       = paste0("Merged interactions – ", names(merged.list)[i])
#     )
#     dev.off()
#   },
#   contentType = "application/pdf"
# )
# 
# # Second plot
# mergedCellchat <- reactiveVal(NULL)
# 
# plotOutput("mergedPlots2", height = "400px")
# 
# # --- Server logic ----
# output$mergedPlots2 <- renderPlot({
#   # ensure inputs are available
#   req(input$levels, 4)
#   
#   # build the grouping factor
#   grp <- factor(
#     rep(input$levels, each = 4),
#     levels = input$levels
#   )
#   
#   # merge interactions on the fly
#   merged.list <- lapply(object.list, function(x) {
#     mergeInteractions(x, grp)
#   })
#   
#   # (re‐)build the combined CellChat if you need it later
#   cellchat.m <<- mergeCellChat(merged.list, add.names = names(merged.list))
#   mergedCellchat(cellchat.m)
#   # compute the max weight for scaling
#   wmax <- getMaxWeight(
#     merged.list,
#     slot.name = c("idents","net","net"),
#     attribute = c("idents","count","count.merged")
#   )
#   
#   netVisual_circle(
#       merged.list[[2]]@net$count.merged,
#       weight.scale     = TRUE,
#       label.edge       = TRUE,
#       edge.weight.max  = wmax[3],
#       edge.width.max   = 12,
#       title.name       = paste0("Merged interactions – ", names(merged.list)[i])
#   )
#   
#   
# })
# 
# downloadButton("download_plots_pdf2", "Download All Plots as PDF")
# 
# output$download_plots_pdf2 <- downloadHandler(
#   filename = function() {
#     paste0("circle_plots_", Sys.Date(), ".pdf")
#   },
#   content = function(file) {
#     req(input$pathway_circle)
#     # open PDF device
#     pdf(file, width = 6, height = 5)
#     # draw one page per pathway
#     netVisual_circle(
#         merged.list[[2]]@net$count.merged,
#         weight.scale     = TRUE,
#         label.edge       = TRUE,
#         edge.weight.max  = wmax[3],
#         edge.width.max   = 12,
#         title.name       = paste0("Merged interactions – ", names(merged.list)[i])
#     )
#     dev.off()
#   },
#   contentType = "application/pdf"
# )

```

##### Merged analysis

We highlight changes in merged interaction networks between conditions.

```{r diff-merged}
# renderPlot({
#   par(mfrow = c(1, 2), xpd = TRUE)
# 
#   netVisual_diffInteraction(
#     mergedCellchat(),
#     weight.scale = TRUE,
#     measure = "count.merged",
#     label.edge = TRUE
#   )
#   
#   netVisual_diffInteraction(
#     mergedCellchat(),
#     weight.scale = TRUE,
#     measure = "weight.merged",
#     label.edge = TRUE
#   )
# })

tagList(
  div(class = "row",
    # Count–merged plot + download
    div(class = "col-md-6",
      plotOutput("mergedDiff_count", height = "400px"),
      tags$br(),
      div(class = "text-center",
        downloadButton("download_diff_count_pdf", "Download PDF"),
        downloadButton("download_diff_count_svg", "Download SVG")
      ),
      style = "margin-bottom: 30px;"
    ),
    # Weight–merged plot + download
    div(class = "col-md-6",
      plotOutput("mergedDiff_weight", height = "400px"),
      tags$br(),
      div(class = "text-center",
        downloadButton("download_diff_weight_pdf", "Download PDF"),
        downloadButton("download_diff_weight_svg", "Download SVG")
      ),
      style = "margin-bottom: 30px;"
    )
  )
)

mergedCellchat <- reactiveVal(NULL)
observe({
  req(input$levels)
  grp <- factor(rep(input$levels, each = 4), levels = input$levels)
  merged.list <- lapply(object.list, mergeInteractions, grp)
  cellchat.m <- mergeCellChat(merged.list, add.names = names(merged.list))
  mergedCellchat(cellchat.m)
})

# 1) Count–merged plot
output$mergedDiff_count <- renderPlot({
  req(mergedCellchat())
  netVisual_diffInteraction(
    mergedCellchat(),
    weight.scale = TRUE,
    measure      = "count.merged",
    label.edge   = TRUE
  )
})
output$download_diff_count_pdf <- downloadHandler(
  filename = function() paste0("diff_count_", Sys.Date(), ".pdf"),
  content = function(file) {
    pdf(file, width = 6, height = 5)
    netVisual_diffInteraction(
      mergedCellchat(),
      weight.scale = TRUE,
      measure      = "count.merged",
      label.edge   = TRUE
    )
    dev.off()
  },
  contentType = "application/pdf"
)

output$download_diff_count_svg <- downloadHandler(
  filename = function() paste0("diff_count_", Sys.Date(), ".svg"),
  content = function(file) {
    svg(file, width = 6, height = 5)
    netVisual_diffInteraction(
      mergedCellchat(),
      weight.scale = TRUE,
      measure      = "count.merged",
      label.edge   = TRUE
    )
    dev.off()
  },
  contentType = NA
)

# 2) Weight–merged plot
output$mergedDiff_weight <- renderPlot({
  req(mergedCellchat())
  netVisual_diffInteraction(
    mergedCellchat(),
    weight.scale = TRUE,
    measure      = "weight.merged",
    label.edge   = TRUE
  )
})

output$download_diff_weight_pdf <- downloadHandler(
  filename = function() paste0("diff_weight_", Sys.Date(), ".pdf"),
  content = function(file) {
    pdf(file, width = 6, height = 5)
    netVisual_diffInteraction(
      mergedCellchat(),
      weight.scale = TRUE,
      measure      = "weight.merged",
      label.edge   = TRUE
    )
    dev.off()
  },
  contentType = "application/pdf"
)

output$download_diff_weight_svg <- downloadHandler(
  filename = function() paste0("diff_weight_", Sys.Date(), ".svg"),
  content = function(file) {
    svg(file, width = 6, height = 5)
    netVisual_diffInteraction(
      mergedCellchat(),
      weight.scale = TRUE,
      measure      = "weight.merged",
      label.edge   = TRUE
    )
    dev.off()
  },
  contentType = NA
)
```

### Compare the major sources and targets in a 2D space {.tabset}

Scatter plots identify major signaling senders and receivers across datasets.

#### Cell populations with significant changes between conditions

```{r signaling-role, fig.height=5, fig.width=10}
# num.link <- sapply(object.list, function(x) {
#   rowSums(x@net$count) + colSums(x@net$count) - diag(x@net$count)
# })
# weight.MinMax <- c(min(num.link), max(num.link))
# gg <- lapply(seq_along(object.list), function(i) {
#   netAnalysis_signalingRole_scatter(
#     object.list[[i]],
#     title = names(object.list)[i],
#     weight.MinMax = weight.MinMax
#   )
# })
# patchwork::wrap_plots(plots = gg)


num.link     <- sapply(object.list, function(x) {
  rowSums(x@net$count) + colSums(x@net$count) - diag(x@net$count)
})
weight.MinMax <- c(min(num.link), max(num.link))
gg <- lapply(seq_along(object.list), function(i) {
  netAnalysis_signalingRole_scatter(
    object.list[[i]],
    title         = names(object.list)[i],
    weight.MinMax = weight.MinMax
  )
})

```

<div style="width:100%; display:flex; justify-content:space-between; align-items:flex-start;">
  <div style="width:48%; text-align:center;">

```{r, scatter-plots-1, fig.width=6, fig.height=5}

gg[[1]]

download_this(
  gg[[1]],
  output_name = paste0("scatter_pdf_", 1),
  output_extension = ".pdf",    
  button_label = "Download PDF",
  button_type = "default",
  self_contained = FALSE,
  ggsave_args = list(width = 4, height = 3, dpi = 300)
)

download_this(
  gg[[1]],
  output_name = paste0("scatter_svg_", 1),
  output_extension = ".svg",    
  button_label = "Download SVG",
  button_type = "default",
  self_contained = FALSE,
  ggsave_args = list(width = 4, height = 3, dpi = 300)
)

```

  </div>

  <div style="width:48%; text-align:center;">
  
```{r, scatter-plots-2, fig.width=6, fig.height=5}

gg[[2]]

download_this(
  gg[[2]],
  output_name = paste0("scatter_pdf_", 2),
  output_extension = ".pdf",    
  button_label = "Download PDF",
  button_type = "default",
  self_contained = FALSE,
  ggsave_args = list(width = 4, height = 3, dpi = 300)
)

download_this(
  gg[[2]],
  output_name = paste0("scatter_svg_", 2),
  output_extension = ".svg",    
  button_label = "Download SVG",
  button_type = "default",
  self_contained = FALSE,
  ggsave_args = list(width = 4, height = 3, dpi = 300)
)

```

  </div>
</div>


<!-- ## Differential Signaling Changes {.tabset} -->

<!-- We assess changes in incoming and outgoing signaling for selected cell types. -->

<!-- ## Inflam. DC -->

<!-- ```{r signaling-changes-dc} -->

<!-- gg1 <- netAnalysis_signalingChanges_scatter( -->

<!--   cellchat, -->

<!--   idents.use = "Inflam. DC", -->

<!--   signaling.exclude = "MIF" -->

<!-- ) -->

<!-- patchwork::wrap_plots(plots = gg1) -->

<!-- ``` -->

<!-- ## cDC1 -->

<!-- ```{r signaling-changes-cdc1} -->

<!-- gg2 <- netAnalysis_signalingChanges_scatter( -->

<!--   cellchat, -->

<!--   idents.use = "cDC1", -->

<!--   signaling.exclude = c("MIF") -->

<!-- ) -->

<!-- patchwork::wrap_plots(plots = gg2) -->

<!-- ``` -->

#### Signalling changes of specific cell populations

```{r signaling-changes-tabs, echo=FALSE}
# 1) UI: selection + placeholder for our tab panel
selectInput(
  inputId  = "idents_use",
  label    = "Select cell types:",
  choices  = levels(cellchat@idents$joint),
  selected = levels(cellchat@idents$joint)[1:3],
  multiple = TRUE
)
uiOutput("signalingTabs")

# 2) Server: build tabs with plot + download buttons
output$signalingTabs <- renderUI({
  req(input$idents_use)
  
  tabs <- lapply(input$idents_use, function(id) {
    safe_id   <- make.names(id)
    plot_name <- paste0("sigPlot_", safe_id)
    png_btn   <- paste0("dl_png_", safe_id)
    pdf_btn   <- paste0("dl_pdf_", safe_id)
    
    tabPanel(
      title = id,
      # plot placeholder
      plotOutput(plot_name, height = "400px"),
      
      # centered download buttons
      tags$div(
        style = "text-align: center; margin-top: 8px;",
        downloadButton(png_btn, "Download PDF"),
        downloadButton(pdf_btn, "Download SVG")
      )
    )
  })
  
  do.call(tabsetPanel, c(id = "tabs", tabs))
})

# 3) Server: render plots and wire up downloadHandlers
observeEvent(input$idents_use, {
  lapply(input$idents_use, function(id) {
    local({
      this_id   <- id
      safe_id   <- make.names(this_id)
      plot_name <- paste0("sigPlot_", safe_id)
      png_btn   <- paste0("dl_png_", safe_id)
      pdf_btn   <- paste0("dl_pdf_", safe_id)
      
      # render the plot
      output[[plot_name]] <- renderPlot({
        netAnalysis_signalingChanges_scatter(
          mergedCellchat(),
          dot.size          = 5,
          label.size        = 4,
          font.size         = 13,
          font.size.title   = 13,
          idents.use        = this_id,
          signaling.exclude = "MIF"
        ) %>% wrap_plots()
      })
      
      # PDF download handler
      output[[pdf_btn]] <- downloadHandler(
        filename = function() paste0("SignalingChanges_", this_id, ".pdf"),
        content  = function(file) {
          ggsave(
            filename = file,
            plot     = netAnalysis_signalingChanges_scatter(
              mergedCellchat(),
              dot.size          = 5,
              label.size        = 4,
              font.size         = 13,
              font.size.title   = 13,
              idents.use        = this_id,
              signaling.exclude = "MIF"
            ) %>% wrap_plots(),
            device   = "pdf",
            width    = 10,
            height   = 6
          )
        }
      )
      
      # PNG download handler
      output[[png_btn]] <- downloadHandler(
        filename = function() paste0("SignalingChanges_", this_id, ".sgv"),
        content  = function(file) {
          ggsave(
            filename = file,
            plot     = netAnalysis_signalingChanges_scatter(
              mergedCellchat(),
              dot.size          = 5,
              label.size        = 4,
              font.size         = 13,
              font.size.title   = 13,
              idents.use        = this_id,
              signaling.exclude = "MIF"
            ) %>% wrap_plots(),
            device   = "sgv",
            dpi      = 300,
            width    = 10,
            height   = 6
          )
        }
      )
      
    })
  })
})
```
